Developing data load and processing

### 2020-07-27 dev/data_preprocessing.Rmd
Work on preprocessing

```{r}
file_relative <- raw_data_map$File[2]
hospital_id <- raw_data_map$HospitalID[2]

@@

@@

patient_data_typed <- patient_data_raw %>% 
  filter(!is.na(Age), !is.na(Sex)) %>%
  transmute(
    hospital_id = hospital_id, 
    age = Age, 
    sex = factor(Sex, levels = c("M","F")), 
    symptom_onset = `Date of symptom onset`, 
    admission = `Date of admission`, 
    transferred_from = `Transferred from`, 
    admitted_for_covid = logical_column(`Admitted for Covid`), 
    ischemic_heart_disease = logical_column(`Ischemic Heart Disease`),
    has_hypertension_drugs = tolower(`Hypertension drugs`) != "no",
    n_hypertension_drugs = if_else(has_hypertension_drugs,
                                   as.integer(if_else(has_hypertension_drugs, `Hypertension drugs`, NA_character_)), #Inner if_else to avoid warning
                                    0L),
    heart_failure = logical_column(`Heart failure`),
    COPD = logical_column(COPD),
    asthma = logical_column(`Asthma Bronchiale`),
    other_lung_disease = logical_column(`Other lung disease`),
    diabetes = logical_column(Diabetes),
    renal_disease = logical_column(`Renal Disease`),
    liver_disease = logical_column(`Liver Disease`),
    smoking = logical_column(Smoking),
    BMI = BMI,
    NYHA = NYHA,
    creatinin = `Creatinin [μmol/L]`,
    pt_inr = `PT INR (Quick)`,
    albumin = `Albumin [g/L]`,
    discontinued_medication = tolower(`Discontinued medication`) != "no",
    discontinued_medication_reason = if_else(discontinued_medication, `Discontinued medication`, NA_character_),
    best_supportive_care_from = if_else(tolower(`Best supportive care from:`) == "no", lubridate::as_date(NA), 
                                        janitor::excel_numeric_to_date(as.numeric(
                                          if_else(tolower(`Best supportive care from:`) == "no", NA_character_, `Best supportive care from:`) # Inner if_else to avoid warning
                                          ))),
    last_record = `Date of last record`,
    outcome = factor(tolower(`Outcome (Discharged/Hospitalized/Transferred/Death)`), levels = tolower(outcome_labels), labels = outcome_labels),
    transferred_to = `Transferred to`
  )
```

>>>>

```{r}
i <- 1
file_relative <- raw_data_map$File[1]
hospital_id <- raw_data_map$HospitalID[1]

@@

patient_data_raw_usable <- patient_data_raw %>% filter(!is.na(Age), !is.na(Sex))

@@

patient_data_typed <- patient_data_raw_usable %>% 
  transmute(
    hospital_id = hospital_id, 
    patient_id = `Patient ID`,
    age = Age, 
    sex = factor(Sex, levels = c("M","F")), 
    symptom_onset = lubridate::as_date(`Date of symptom onset`), 
    admission = lubridate::as_date(`Date of admission`), 
    transferred_from = `Transferred from`, 
    admitted_for_covid = logical_column(`Admitted for Covid`), 
    ischemic_heart_disease = logical_column(`Ischemic Heart Disease`),
    has_hypertension_drugs = tolower(`Hypertension drugs`) != "no",
    n_hypertension_drugs = if_else(has_hypertension_drugs,
                                   as.integer(if_else(has_hypertension_drugs, `Hypertension drugs`, NA_character_)), #Inner if_else to avoid warning
                                    0L),
    heart_failure = logical_column(`Heart failure`),
    COPD = logical_column(COPD),
    asthma = logical_column(`Asthma Bronchiale`),
    other_lung_disease = logical_column(`Other lung disease`),
    diabetes = logical_column(Diabetes),
    renal_disease = logical_column(`Renal Disease`),
    liver_disease = logical_column(`Liver Disease`),
    smoking = logical_column(Smoking),
    BMI = BMI,
    NYHA = NYHA,
    creatinin = `Creatinin [μmol/L]`,
    pt_inr = `PT INR (Quick)`,
    albumin = `Albumin [g/L]`,
    discontinued_medication = tolower(`Discontinued medication`) != "no",
    discontinued_medication_reason = if_else(discontinued_medication, `Discontinued medication`, NA_character_),
    best_supportive_care_from = if_else(tolower(`Best supportive care from:`) == "no", lubridate::as_date(NA), 
                                        janitor::excel_numeric_to_date(as.numeric(
                                          if_else(tolower(`Best supportive care from:`) == "no", NA_character_, `Best supportive care from:`) # Inner if_else to avoid warning
                                          ))),
    last_record = lubridate::as_date(`Date of last record`),
    outcome = factor(tolower(`Outcome (Discharged/Hospitalized/Transferred/Death)`), levels = tolower(outcome_labels), labels = outcome_labels),
    transferred_to = `Transferred to`
  )

invalid_outcome <- is.na(patient_data_raw_usable$`Outcome (Discharged/Hospitalized/Transferred/Death)`) != is.na(patient_data_typed$outcome)
if(any(invalid_outcome)) {
  print(patient_data_raw_usable %>% filter(invalid_outcome))
  stop("Invalid outcome")
}
```

```{r}
progression_data_raw <- readxl::read_excel(file, sheet = "Disease Progression", range = "A1:CU2539", na = c("NA",""), col_types = c(c("text","text", "skip", "text", "text"), rep("text", 94)), col_names = TRUE)

progression_data <- progression_data_raw %>% 
  rename(patient_id = `Patient ID`, indicator = Indicator, first_day = First_Day, unit = Unit) %>%
  filter(!is.na(indicator))

# Remove the "Don't edit examples" comment
progression_data$patient_id[2:9] <- NA_character_

# Spread Patient ID and Start date
last_patient_id <- progression_data$patient_id[1]
last_first_day <- progression_data$first_day[1]
for(i in 2:nrow(progression_data)) {
  if(!is.na(progression_data$patient_id[i])) {
    last_patient_id <- progression_data$patient_id[i]
    last_first_day <- progression_data$first_day[i]
  } else {
    progression_data$patient_id[i] <- last_patient_id
    progression_data$first_day[i] <- last_first_day
  }
}

progression_data <- progression_data %>% filter(patient_id %in% patient_data_typed$patient_id)

progression_data <- progression_data %>% 
  pivot_longer(starts_with("Day_"), names_to = "day", values_to = "value", names_prefix = "Day_") %>%
  mutate(day = as.integer(day))

progression_data <- progression_data %>% 
  filter(!is.na(indicator))

# Check dates
dates_data <- progression_data %>% 
  filter(indicator == "Date")
test_dates <- dates_data %>%
  mutate(value = as.integer(value), first_day = as.integer(first_day)) %>%
  filter(value != first_day + day - 1 )

if(nrow(test_dates) > 0) {
  stop("Dates check failed")
}

breathing_data <- progression_data %>% filter(indicator == "Breathing") %>%
  mutate(breathing = factor(tolower(value), levels = tolower(c("AA","Oxygen", "NIPPV","MV","ECMO")),labels = c("AA","Oxygen", "NIPPV","MV","ECMO"), ordered = TRUE))

breathing_unrecognized <- breathing_data %>% filter(is.na(breathing) != is.na(value))
if(nrow(breathing_unrecognized) > 0) {
  stop("Breathing NAs check failed")
}

breathing_data <- breathing_data %>% select(-value, -indicator, -unit)

correct_first_day_admission <- function(patient_data, progression_data) {
  nrow_before <- nrow(progression_data)
  progression_data <- progression_data %>% 
    inner_join(patient_data %>% select(patient_id, admission), by = "patient_id") %>%
    mutate(first_day = janitor::excel_numeric_to_date(as.numeric(first_day)),
           correction = as.integer(first_day - admission),
           day = day + correction - 1)

  if(nrow(progression_data) != nrow_before) {
    stop("Invalid join")
  }

  progression_data %>%
    select(-admission, -first_day, -correction)
}

breathing_data <- correct_first_day_admission(patient_data_typed, breathing_data) %>% filter(!is.na(breathing))

adverse_event_data <- progression_data %>% 
  filter(indicator == "Adverse events")

dummy_markers <- c("Drugs", "Other marker (optional)", "Other Covid drug", "Yet another drug")

marker_data <- progression_data %>% filter(!(indicator %in% c("Date", "Breathing", "Adverse events", dummy_markers))) %>% rename(marker = indicator)
marker_data <- correct_first_day_admission(patient_data_typed, marker_data)

# Correct Excel error
marker_data <- marker_data %>%
  mutate(marker = case_when(
    grepl("^Supp\\. O[0-9]*$", marker) ~ "Supp. O2",
    grepl("^SpO[0-9]*$", marker) ~ "SpO2",
    grepl("^IL-[0-9]*$", marker) ~ "IL-6",
    TRUE ~ marker))

if(any(is.na(marker_data$marker))) {
  stop("NA in markers")
}

marker_map <- c("PCR" = "pcr_value", 
                "Supp. O2" = "oxygen_flow",
                "SpO2" = "SpO2",
                "SpO2 native" = "SpO2_native",
                "Horowitz index" = "Horowitz_index",
                "PEEP" = "PEEP",
                "Ferritin" = "ferritin",
                "D-dimer" = "d_dimer",
                "CRP" = "CRP",
                "IL-6" = "IL_6",
                "Lymphocyte count" = "lymphocyte_count",
                "Hydroxychloroquine" = "hcq",
                "Azithromycin" = "az",
                "Kaletra" = "kaletra",
                "Tocilizumab" = "tocilizumab",
                "Amoxicillin/Clavulanate" = "amoxiclav",
                "Piperacilin/tazobactam" = "piperacilin_tazobactam",
                "Clarithromycin" = "clarithromycin",
                "ofloxacin" = "ofloxacin",
                "ceftriaxone" = "ceftriaxone",
                "cefuroxim" = "cefuroxim",
                "ampicillin/sulbactam" = "ampicillin_sulbactam"
                )
#Check NAs (TODO)
unrecgonized_markers <- marker_data %>% filter(!(tolower(marker) %in% tolower(names(marker_map))))

if(nrow(unrecgonized_markers) > 0) {
  print(unrecgonized_markers)
  stop("Unrecognized markers")
}

names(marker_map) <- tolower(names(marker_map))

marker_data <- marker_data %>% mutate(marker = marker_map[tolower(marker)])

#TODO - handle PCR specifically

```
